A meta-analysis of mental rotation ability in the first years of life

### 2022-05-05 code/analysis.Rmd
Update intro, main analysis

---
title             : "A small male advantage in infant mental rotation"
shorttitle        : "Infant mental rotation"

author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
    corresponding : "yes"
    address       : "Stephanstraße 1a, 04103 Leipzig"
    email         : "enge@cbs.mpg.de"
  - name          : "Shreya Kapoor"
    affiliation   : "1"
  - name          : "Anne-Sophie Kieslinger"
    affiliation   : "1"
  - name          : "Michael A. Skeide"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Research Group Learning in Early Childhood, Max Planck Institute for Human Cognitive and Brain Sciences"
  - id            : "2"
    institution   : "Department of Psychology, Humboldt-Universität zu Berlin"

@@

abstract: |
  Mental rotation, the cognitive process of moving an object in mind to predict how it looks in a new orientation, is tightly coupled to intelligence, learning and educational achievement (Shepard & Metzler, 1971; Hegarty & Kozhevnikov, 1999; Johnshon & Bouchard, 2005). On average, males solve mental rotation tasks slightly faster than females (Linn & Petersen 1985, Voyer et al. 1995, Voyer 2011, Maeda & Yoon 2013). When such behavioral differences emerge during development, however, remains poorly understood (Lauer et al. 2019). Here we analyzed effect sizes derived from 59 experiments conducted in 1,798 infants aged 3 to 16 months. We robustly found that male infants looked longer at novel rotated objects than female infants. The effect size of this average difference was small and did not change with age. These findings indicate that gender differences in mental rotation are already present in the first months of life.

@@

bibliography      : ["r-references.bib"]
csl               : "nature.csl"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

@@

```{r, setup, include=FALSE}
# Load packages
library(papaja)
library(here)
library(ggridges)
library(tidyverse)
library(magrittr)
library(metafor)
library(brms)
library(bayestestR)
library(tidybayes)
library(cowplot)

@@

# We go ahead with the last solution as to use the maximum amount of information
# while not biasing the gender results
dat <- dat_both_mixed
```

@@

The cognitive ability to move visual object representations in mind for recognition across different orientations, known as mental rotation, emerges in the first three months of life (Johnson & Moore 2020, Moore & Johnson 2020). Mental rotation is a key component of intelligence and a powerful predictor of learning outcome and educational achievement (Shepard & Metzler, 1971; Hegarty & Kozhevnikov, 1999; Johnshon & Bouchard, 2005).

Previous meta-analyses revealed that males solve mental rotation tasks slightly faster than females on average (Linn & Petersen 1985, Voyer et al. 1995, Voyer 2011, Maeda & Yoon 2013). Effect sizes of this difference, however, are heterogeneous and often only medium (mean weighted g = 0.37–0.73). Interestingly, a recent meta-analysis in 3-17-year-old children and adolescents suggests even only a small-to-medium difference (mean weighted g = 0.39) (Lauer et al. 2019). Whether gender differences in mental rotation behavior already emerge during infancy remains unknown.

In the present study we meta-analyzed 59 effect sizes derived from looking times in mental rotation tasks conducted by 1,798 infants (47.5% female) aged 3 to 16 months (mean age of 7 months 14 days). All tasks were embedded either into habituation experiments (40) or violation of expectation experiments (19) (Fantz 1964; Baillargeon et al. 1985). These experiments comprised real world stimuli (e.g., toy objects, Antrilli & Wang, 2016), three-dimensional digital stimuli (e.g., cube figures; Moore & Johnson, 2008), or two-dimensional digital stimuli (e.g., digits; Quinn & Liben, 2008). In habituation experiments, infants repeatedly saw an object until their looking times declined before they were presented with a mirror image of the object. Longer looking times at the mirror image were taken as evidence that an infant still recognized the familiar object after rotation through the new orientation. In violation of expectation experiments, infants were habituated to an object that was revolving repeatedly through a certain angle. Then, either the familiar object or an unseen object was shown while they were revolving through a certain angle. Subsequently, the object disappeared behind an occluder. Finally, the occluder was removed and either the object or its mirror object were shown. Larger differences in looking times for the familiar object versus the unseen mirror image (at the new angle) were taken as evidence that an infant still recognized the familiar object after rotation through the new angle. In contrast, looking times for both objects were expected to be similar if an infant did not recognize the familiar object from the new angle.
Following previous work in older populations we hypothesized that looking times in mental rotation experiments are on average significantly longer in male compared to female infants. Effect sizes were assumed to be small.

@@

# Run prior predictive simulation
n_iter <- 20000
n_warmup <- 1000
n_chains <- 4
res_prior <- brm(
  gi | se(sei) ~ 0 + Intercept + (1 | article / experiment),
  data = dat,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = n_chains,
  iter = n_iter,
  warmup = n_warmup,
  cores = n_chains,
  control = list(adapt_delta = 0.99),
  file = here("results", "models", "res_prior")
)
summary(res_prior)
# plot(res_prior)

@@

@@

# Compute by-experiment variance
var_experiment <- as.matrix(
  res_brm,
  variable = "sd_article:experiment__Intercept"
)^2
mean_qi(var_experiment)

# Compute by-article variance
var_article <- as.matrix(
  res_brm,
  variable = "sd_article__Intercept"
)^2
mean_qi(var_article)

# Compute intra-class correlation
var_total <- (var_experiment + var_article)
icc_draws <- var_article / var_total
mean_qi(icc_draws)
```

For our first meta-analytic model effect sizes were quantified as the standardized mean difference in infants’ looking times for novel and familiar rotated objects. Using this effect size index we ran a Bayesian three-level random-effects model to test if there was evidence that infants did perform mental rotation. Across studies, infants indeed looked longer at novel rotated objects than at familiar rotated objects, with a standardized mean difference of g = 0.26, 95% credible interval (CrI) [0.12, 0.41] (Fig. 1). The probability for this effect being greater than zero was 99.9%. The heterogeneity of effect sizes was $sigma$2Experiment = 0.03, 95% CrI [0.00, 0.12], at the experiment level and $sigma$2Article = 0.17, 95% CrI [0.09, 0.29] at the article level. Therefore, approximately 86.7% of the heterogeneity between effect sizes was attributable to differences between experiments and 13.3% was attributable to differences between articles.

```{r, fig1, include=TRUE, fig.height=14, fig.width=12, fig.cap="(ref:fig1-caption)"}
# Helper functions for printing confidence intervals
print_num <- function(x, digits = 2) {
  x <- format(round(x, digits), trim = TRUE, nsmall = digits)
  x[x == "NA"] <- NA
  return(x)
}
print_ci <- function(est, lb, ub) {
  paste0(print_num(est), " [", print_num(lb), ", ", print_num(ub), "]")
}

# Get posterior draws for each experiment
draws <- epred_draws(res_brm, dat, ndraws = NULL) %>%
  mutate(experiment_f = factor(experiment))

# Get posterior draws for the meta-analytic effect
meta_draws <- spread_draws(res_brm, b_Intercept, ndraws = NULL)
meta_effect_size <- summary(res_brm)$fixed

# Create forest plot
n_experiments <- length(unique(dat$experiment))
dat %>%
  # Make sure the plot will be ordered alphabetically by experiment IDs
  arrange(experiment) %>%
  mutate(
    experiment_f = fct_rev(fct_expand(factor(experiment), "model")),
    # Show article labels only for the first experiment per article
    article = if_else(article == lag(article, default = ""), "", article),
    # Compute frequentist confidence intervals
    ci_lb = gi - qnorm(0.05 / 2, lower.tail = FALSE) * sqrt(vi),
    ci_ub = gi + qnorm(0.05 / 2, lower.tail = FALSE) * sqrt(vi),
    ci_print = print_ci(gi, ci_lb, ci_ub)
  ) %>%
  # Prepare plotting canvas
  ggplot(aes(x = gi, y = experiment_f)) +
  geom_vline(xintercept = seq(-3, 3, 0.5), color = "grey90") +
  # Add article and group labels as text on the left
  geom_text(aes(x = -9.9, label = article), hjust = 0) +
  geom_text(aes(x = -7.1, label = group_long), hjust = 0) +
  # Add experiment-specific effect sizes and CIs as text on the right
  geom_text(aes(x = 3.7, label = print_num(gi)), hjust = 1) +
  geom_text(aes(x = 4.4, label = print_num(ci_lb)), hjust = 1) +
  geom_text(aes(x = 5.1, label = print_num(ci_ub)), hjust = 1) +
  # Add Bayesian highest posterior density intervals for each experiment
  stat_interval(
    aes(x = .epred),
    data = draws,
    alpha = .8,
    point_interval = "mean_qi",
    .width = c(0.5, 0.95),
  ) +
  scale_color_grey(
    start = 0.85, end = 0.65,
    labels = as_mapper(~ scales::percent(as.numeric(.x)))
  ) +
  # Add experiment-specific effect sizes and CIs as dots with error bars
  geom_linerange(aes(xmin = ci_lb, xmax = ci_ub), size = 0.35) +
  geom_point(aes(size = ni), shape = 22, fill = "white") + # Or (1 / vi)?
  # Add posterior distribution for the meta-analytic effect
  stat_halfeye(
    aes(x = b_Intercept, y = -1),
    meta_draws,
    point_interval = "mean_qi",
    .width = c(0.95)
  ) +
  annotate(
    "text",
    x = c(-9.9, 3.7, 4.4, 5.1),
    y = -0.5,
    label = c(
      "Three-level model",
      print_num(meta_effect_size$Estimate),
      print_num(meta_effect_size$`l-95% CI`),
      print_num(meta_effect_size$`u-95% CI`)
    ),
    hjust = c(0, 1, 1, 1),
    fontface = "bold"
  ) +
  # Add headers for each of the for columns
  annotate( # Just to make the grid lines disappear
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = n_experiments + 0.5,
    ymax = Inf,
    fill = "white"
  ) +
  annotate(
    "text",
    x = c(-9.9, -7.1, 0, 3.7, 4.4, 5.1),
    y = n_experiments + 1.5,
    label = c(
      "Article",
      "Experiment",
      "Effect size plot",
      "Effect size",
      "Lower",
      "Upper"
    ),
    hjust = c(0, 0, 0.5, 1, 1, 1),
    fontface = "bold"
  ) +
  # Styling
  coord_cartesian(
    ylim = c(-1.5, n_experiments + 2), expand = FALSE, clip = "off"
  ) +
  scale_x_continuous(breaks = seq(-3, 3, 0.5)) +
  annotate("segment", x = -3.3, xend = 3.3, y = -1.5, yend = -1.5) +
  labs(
    x = expression("Hedges'" ~ italic("g")),
    size = "Sample size",
    color = "CrI level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(family = "Helvetica", color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.x = element_line(colour = "black"),
    axis.title.x = element_text(hjust = 0.67, margin = margin(b = -15)),
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    text = element_text(family = "Helvetica")
  )

@@

(ref:fig1-caption) Mental rotation performance. A Bayesian three-level meta-analysis provided evidence for mental rotation ability in infants. White squares depict the effect sizes (Hedges’ g) for infants’ mental rotation performance in all individual experiments and black lines depict their 95% confidence intervals. Gray bars indicate the 95% and 50% Bayesian credible intervals (CrI). These intervals got shrunk towards the meta-analytic effect size because of partial pooling which regularized the impact of experiments with small sample sizes or unrealistically large effect sizes to prevent overfitting the model. The last line shows the meta-analytic effect size (black dot) together with its 95% CrI and its posterior distribution (gray curve).

@@

# Compute intra-class correlation
var_total_reg <- var_experiment_reg + var_article_reg
icc_reg <- var_article_reg / var_total_reg
mean_qi(icc_reg)

# Compute reduction in heterogeneity compared to the original meta-analysis
1 - mean(var_total_reg / var_total)
```

@@

(ref:fig2-caption) Effects of gender, age, and task. a Squares show the effect size (Hedges’ g) for infants’ mental rotation performance in each of the 59 individual experiments. Squares are color-coded according to the type of habituation task and the gender of the infants (blue = habituation task, all-female sample, yellow = habituation task, all-male sample, red = habituation task, mixed-gender sample, purple = violation of expectation [VoE] task, mixed-gender sample). Lines indicate the best-fit regression estimates according to a Bayesian three-level meta-regression model and gray ribbons indicate their corresponding 95% and 50% credible interval (CrI). b Fixed effect estimates obtained from the Bayesian three-level meta-regression model are depicted as black dots together with their 95% Crl (thin black lines) and 50% Crl (thick black lines). Gray curves indicate the posterior distribution for each effect.

To confirm the gender difference between males and females we set up another Bayesian meta-analysis, this time focusing on the looking-time contrasts between all-male and all-female groups reported within each article by the original authors. Our additional analysis revealed a meta-analytic effect size of g = 0.20, 95% Crl [0.03, 0.38] (Supplementary Fig. 1) and a probability of this effect being greater than zero of 98.9%. The heterogeneity of effect sizes was $sigma$2Experiment = 0.01, 95% CrI [0.00, 0.07], at the experiment level, and $sigma$2Article = 0.02, 95% CrI [0.00, 0.10] at the article level. This means that 56.1% of the total variance can be attributed to differences between articles and 43.1% to differences within articles.

@@

```{r, fig4, include=TRUE, fig.width=6, fig.cap="(ref:fig4-caption)", out.width="50%"}
# Include the flowchart (created with MS PowerPoint)
knitr::include_graphics("results/figures/flowchart.pdf")
```

@@

We entered the search terms (“mental rotation” OR “mental transformation” OR “spatial rotation” OR “spatial transformation” OR “spatial ability” OR “spatial skills”) AND (“infant” OR “infants” OR “infanthood” OR “toddler” OR “toddlers” OR “toddlerhood” OR “child” OR “children” OR “childhood” OR “month” OR “months”) into four online databases (APA PsycINFO, PubMed/MEDLINE, Scopus, ProQuest Dissertations & Theses Global). All database queries were completed on December 6, 2021. We configured the databases to check for article titles, abstracts, and keywords while applying no other filters or limits. This yielded 2,616 articles in total, 1,954 of which remained after removing duplicate records (Figure 1). We further identified 76 articles by screening the reference sections of previous reviews and meta-analyses on mental rotation and related skills (Frick et al., 2014; Johnson & Moore, 2020; Kubicek & Schwarzer, 2018; Lauer et al., 2019; Linn & Petersen, 1985; Moore & Johnson, 2020; Uttal et al., 2013; Voyer et al., 1995; Yang et al., 2020). Of these, 34 had not been covered by the database search. We also identified 94 articles by screening the reference sections of all articles that had been included after the first pass of the selection process. Of these, 49 had not been covered by the database search. We thus screened 2,037 unique articles in total.

@@

Two independent raters read the abstract and, if necessary, relevant sections of the full text to check if an article fulfilled the inclusion criteria. Interrater agreement for the binary decision to include versus exclude an article was 98.5% ($Kappa$w [Cohen’s weighted kappa] = 0.67, 95% CI [0.55, 0.78]). Interrater agreement for the specific eligibility criteria were 88.2% ($Kappa$w = 0.72, 95% CI [0.40, 1.00]). Cases where the two ratings diverged were resolved via discussion among all raters until a consensus was reached. One article (Mash et al., 2007) was excluded because the authors used a unique mental rotation paradigm that was not comparable to that used in the other articles. One article (Pedrett et al., 2020) was excluded because the average age of the infants (30.7 months) was almost twice as high as that of the next article (15.8 months), z = ??? compared to all articles. We therefore decided to narrow our analysis from the first 3 years of life to the first 16 months of life. This procedure led to a total of 23 articles being included in the meta-analysis (Fig. 4).

Many of these articles consisted of multiple experiments, e.g., using different variations of the mental rotation task or different subpopulations of infants. We included all of these experiments in the meta-analysis and accounted for the dependencies between them by means of multilevel modeling with by-article random effects (see below). However, we excluded experiments if there was insufficient information to compute a standardized effect size (see below). Whenever an article reported separate effect sizes for boys and girls—or other subgroups like crawling and non-crawling infants)—but also an effect size combining these groups, we only included the combined effect size. We dropped effect sizes that were clearly based on the same data but reported in different articles. This procedure led to a total of 59 experiments being included in the meta-analysis (Supplementary Table 1).

@@

One outcome measure per experiment was converted into a standardized mean difference with small sample correction (Hedges’ g) using the formulas provided in Table 1. The sampling variance Var_g of the outcome measure g for each experiment was computed using the formula provided by Goulet-Pelletier & Cousineau (2018):

v = (df(df - 2)) * ((2 * (1 - r)) / n) * (1 + g^2 * (n / (2 * (1 - r)))) - (g^2 / j^2)

>>>>

---
title             : "Gender differences in infant mental rotation"
shorttitle        : "Infant mental rotation"

author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2,*"
    corresponding : "yes"
    address       : "Stephanstraße 1a, 04103 Leipzig"
    email         : "enge@cbs.mpg.de"
  - name          : "Shreya Kapoor"
    affiliation   : "1,*"
  - name          : "Anne-Sophie Kieslinger"
    affiliation   : "1,*"
  - name          : "Michael A. Skeide"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Research Group Learning in Early Childhood, Max Planck Institute for Human Cognitive and Brain Sciences"
  - id            : "2"
    institution   : "Department of Psychology, Humboldt-Universität zu Berlin"
  - id            : "*"
    institution   : "These authors contributed equally"

@@

abstract: |
  Mental rotation, the cognitive process of moving an object in mind to predict how it looks in a new orientation, is tightly coupled to intelligence, learning and educational achievement [@shepard1971; @hegarty1999; @johnson2005]. On average, males solve mental rotation tasks slightly faster than females [@linn1985; @voyer1995; @voyer2011; @maeda2013]. When such behavioral differences emerge during development, however, remains poorly understood [@lauer2019a]. Here we analyzed effect sizes derived from 59 experiments conducted in 1,798 infants aged 3 to 16 months. We robustly found that male infants looked longer at novel rotated objects than female infants. The effect size of this average difference was small and did not change with age. These findings indicate that gender differences in mental rotation are already present in the first months of life.

@@

bibliography      : "references.bib"
csl               : "nature.csl"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

@@

```{r, setup, include=FALSE}
# Load packages
library(papaja)
library(here)
library(scales)
library(tidyverse)
library(magrittr)
library(metafor)
library(brms)
library(bayestestR)
library(tidybayes)
library(cowplot)

@@

# We go ahead with the last solution as to use the maximum amount of information
# while not biasing the gender results
dat <- dat_both_mixed

# Helper functions to re-format age from days to months + days
format_days_months <- function(days, long = FALSE) {
  m_str <- ifelse(long, " months ", "m ")
  d_str <- ifelse(long, " days", "d")
  days_per_month <- 30.417
  months <- days %/% days_per_month
  days_left <- as.integer(days - months * days_per_month)
  str <- paste0(as.character(months), m_str, as.character(days_left), d_str)
  str[is.na(days)] <- NA
  return(str)
}

# Helper function for printing percentages
print_perc <- function(x) scales::percent(x, accuracy = 0.1)

# Extract some descriptive statistics for usage in the main text
descs <- list(
  n_experiments = nrow(dat),
  n_infants = sum(dat$sample_size),
  percent_female = mean(dat$female_percent, na.rm = TRUE),
  min_age_months = as.integer(min(dat$age_min, na.rm = TRUE) / 30.417),
  mean_age_weighted = format_days_months(
    sum(dat$age_mean * dat$ni) / (sum(dat$ni)),
    long = TRUE
  ),
  n_experiments_habituation = nrow(filter(dat, task == "Habituation")),
  n_experiments_voe = nrow(filter(dat, task == "VoE"))
)
```

@@

The cognitive ability to move visual object representations in mind for recognition across different orientations, known as mental rotation, emerges in the first three months of life [@moore2020; @johnson2020]. Mental rotation is a key component of intelligence and a powerful predictor of learning outcome and educational achievement [@shepard1971; @hegarty1999; @johnson2005].

Previous meta-analyses revealed that males solve mental rotation tasks slightly faster than females on average [@linn1985; @voyer1995; @voyer2011; @maeda2013]. Effect sizes of this difference, however, are heterogeneous and often only medium in size (mean weighted $g = 0.37\text{--}0.73$). Interestingly, a recent meta-analysis [@lauer2019a] in 3-17-year-old children and adolescents suggests even only a small-to-medium difference (mean weighted $g = 0.39$). Whether gender differences in mental rotation behavior already emerge during infancy remains unknown.

In the present study, we meta-analyzed `r descs$n_experiments` effect sizes derived from looking times in mental rotation tasks conducted by `r descs$n_infants` infants (`r print_perc(descs$percent_female)` female) aged `r descs$min_age_months` to `r descs$max_age_months` months (mean age of `r descs$mean_age_weighted`; Supplementary Table 1). All tasks were embedded either into habituation experiments (`r descs$n_experiments_habituation`) or violation of expectation experiments (`r descs$n_experiments_voe`) [@fantz1964; @baillargeon1985]. These experiments comprised real world stimuli (e.g., toy objects [@antrilli2016]), three-dimensional digital stimuli (e.g., cube figures [@moore2008]), or two-dimensional digital stimuli (e.g., digits [@quinn2008]). In habituation experiments, infants repeatedly saw an object until their looking times declined before they were presented with a mirror image of the object. Longer looking times at the mirror image were taken as evidence that an infant still recognized the familiar object after rotation through the new orientation. In violation of expectation experiments, infants were habituated to an object that was revolving repeatedly through a certain angle. Then, either the familiar object or an unseen object was shown while they were revolving through a certain angle. Subsequently, the object disappeared behind an occluder. Finally, the occluder was removed and either the object or its mirror object were shown. Larger differences in looking times for the familiar object versus the unseen mirror image (at the new angle) were taken as evidence that an infant still recognized the familiar object after rotation through the new angle. In contrast, looking times for both objects were expected to be similar if an infant did not recognize the familiar object from the new angle.

Following previous work in older populations, we hypothesized that looking times in mental rotation experiments are on average significantly longer in male compared to female infants. Effect sizes were assumed to be small.

@@

# Run prior predictive simulation
n_iter <- 4000
n_warmup <- 1000
n_chains <- 4
res_prior <- brm(
  gi | se(sei) ~ 0 + Intercept + (1 | article / experiment),
  data = dat,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = n_chains,
  iter = n_iter,
  warmup = n_warmup,
  cores = n_chains,
  control = list(adapt_delta = 0.99),
  file = here("results", "models", "res_prior")
)
summary(res_prior)
# plot(res_prior)

@@

# Get posterior probability mass > 0 for the meta-analytic effect
(hyp_brm <- hypothesis(res_brm, "Intercept > 0")$hypothesis)

@@

# Extract posterior draws for the meta-analytic effects
spread_draws(
  res_brm,
  b_Intercept, sd_article__Intercept, `sd_article:experiment__Intercept`,
  ndraws = NULL
) %>%
  # Compute variances and ICC
  mutate(
    intercept = b_Intercept,
    sigma_article = sd_article__Intercept,
    sigma_experiment = `sd_article:experiment__Intercept`,
    sigma2_article = sigma_article^2,
    sigma2_experiment = sigma_experiment^2,
    sigma2_total = sigma2_article + sigma2_experiment,
    icc = sigma2_article / sigma2_total,
    .keep = "unused"
  ) -> draws_brm

# Summarize as mean and 95% credible intervals
(summ_brm <- as.list(mean_qi(draws_brm)))
```

```{r, fig1, include=TRUE, fig.height=14, fig.width=12, fig.cap="(ref:fig1-caption)"}
# Helper functions for printing numbers with exactly 2 decimal points
print_num <- function(x, digits = 2) {
  x <- format(round(x, digits), trim = TRUE, nsmall = digits)
  x[x == "NA"] <- NA
  return(x)
}

# Helper function for confidence intervals (optionally adding the mean)
print_ci <- function(lb, ub, mean = NULL) {
  mean_str <- ifelse(mean, paste0(print_num(mean), " "))
  paste0(mean_str, "[", print_num(lb), ", ", print_num(ub), "]")
}

# Get posterior draws for the effect in each experiment
epred_draws_brm <- epred_draws(res_brm, dat, ndraws = NULL) %>%
  mutate(experiment_f = factor(experiment))

# Create forest plot
dat %>%
  # Make sure the plot will be ordered alphabetically by experiment IDs
  arrange(experiment) %>%
  mutate(
    experiment_f = fct_rev(fct_expand(factor(experiment), "model")),
    # Show article labels only for the first experiment per article
    article = if_else(article == lag(article, default = ""), "", article),
    # Compute frequentist confidence intervals
    ci_lb = gi - qnorm(0.05 / 2, lower.tail = FALSE) * sqrt(vi),
    ci_ub = gi + qnorm(0.05 / 2, lower.tail = FALSE) * sqrt(vi),
    ci_print = print_ci(ci_lb, ci_ub, mean = gi)
  ) %>%
  # Prepare plotting canvas
  ggplot(aes(x = gi, y = experiment_f)) +
  geom_vline(xintercept = seq(-3, 3, 0.5), color = "grey90") +
  # Add article and group labels as text on the left
  geom_text(aes(x = -9.9, label = article), hjust = 0) +
  geom_text(aes(x = -7.1, label = group_long), hjust = 0) +
  # Add experiment-specific effect sizes and CIs as text on the right
  geom_text(aes(x = 3.7, label = print_num(gi)), hjust = 1) +
  geom_text(aes(x = 4.4, label = print_num(ci_lb)), hjust = 1) +
  geom_text(aes(x = 5.1, label = print_num(ci_ub)), hjust = 1) +
  # Add Bayesian highest posterior density intervals for each experiment
  stat_interval(
    aes(x = .epred),
    data = epred_draws_brm,
    alpha = .8,
    point_interval = "mean_qi",
    .width = c(0.5, 0.95),
  ) +
  scale_color_grey(
    start = 0.85, end = 0.65,
    labels = as_mapper(~ scales::percent(as.numeric(.x)))
  ) +
  # Add experiment-specific effect sizes and CIs as dots with error bars
  geom_linerange(aes(xmin = ci_lb, xmax = ci_ub), size = 0.35) +
  geom_point(aes(size = ni), shape = 22, fill = "white") + # Or (1 / vi)?
  # Add posterior distribution for the meta-analytic effect
  stat_halfeye(
    aes(x = intercept, y = -1),
    draws_brm,
    point_interval = "mean_qi",
    .width = c(0.95)
  ) +
  annotate(
    "text",
    x = c(-9.9, 3.7, 4.4, 5.1),
    y = -0.5,
    label = c(
      "Three-level model",
      print_num(summ_brm$intercept),
      print_num(summ_brm$intercept.lower),
      print_num(summ_brm$intercept.upper)
    ),
    hjust = c(0, 1, 1, 1),
    fontface = "bold"
  ) +
  # Add headers for each of the for columns
  annotate(
    "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = descs$n_experiments + 0.5,
    ymax = Inf,
    fill = "white"
  ) +
  annotate(
    "text",
    x = c(-9.9, -7.1, 0, 3.7, 4.4, 5.1),
    y = descs$n_experiments + 1.5,
    label = c(
      "Article",
      "Experiment",
      "Effect size plot",
      "Effect size",
      "Lower",
      "Upper"
    ),
    hjust = c(0, 0, 0.5, 1, 1, 1),
    fontface = "bold"
  ) +
  # Styling
  coord_cartesian(
    ylim = c(-1.5, descs$n_experiments + 2), expand = FALSE, clip = "off"
  ) +
  scale_x_continuous(breaks = seq(-3, 3, 0.5)) +
  annotate("segment", x = -3.3, xend = 3.3, y = -1.5, yend = -1.5) +
  labs(
    x = expression("Hedges'" ~ italic("g")),
    size = "Sample size",
    color = "CrI level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(family = "Helvetica", color = "black"),
    axis.text.y = element_blank(),
    axis.ticks.x = element_line(colour = "black"),
    axis.title.x = element_text(hjust = 0.67, margin = margin(b = -15)),
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    text = element_text(family = "Helvetica")
  )

@@

(ref:fig1-caption) Mental rotation performance. A Bayesian three-level meta-analysis provided evidence for mental rotation ability in infants. White squares depict the effect sizes (Hedges’ g) for infants’ mental rotation performance in all individual experiments and black lines depict their 95% confidence intervals. Gray bars indicate the 95% and 50% Bayesian credible intervals (CrI). These intervals got shrunk towards the meta-analytic effect size because of partial pooling which regularized the impact of experiments with small sample sizes or unrealistically large effect sizes to prevent overfitting the model. The last line shows the meta-analytic effect size (black dot) together with its 95% CrI (black line) and its posterior distribution (gray curve).

For our first meta-analytic model effect sizes were quantified as the standardized mean difference in infants’ looking times for novel and familiar rotated objects. Using this effect size index we ran a Bayesian three-level random-effects model to test if there was evidence that infants did perform mental rotation. Across studies, infants indeed looked longer at novel rotated objects than at familiar rotated objects, with a standardized mean difference of $g = `r summ_brm$intercept`$, 95% credible interval (CrI) [`r summ_brm$intercept.lower`, `r summ_brm$intercept.upper`] (Fig. 1). The probability for this effect being greater than zero was `r print_perc(hyp_brm$Post.Prob)`. The heterogeneity of effect sizes was $\sigma_{\text{Experiment}}^2 = `r summ_brm$sigma2_experiment`$, 95% CrI [`r summ_brm$sigma2_experiment.lower`, `r summ_brm$sigma2_experiment.upper`], at the experiment level and $\sigma_{\text{Article}}^2 = `r summ_brm$sigma2_article`$, 95% CrI [`r summ_brm$sigma2_article.lower`, `r summ_brm$sigma2_article.upper`] at the article level. Therefore, approximately `r print_perc(1 - summ_brm$icc)` of the heterogeneity between effect sizes was attributable to differences between experiments and `r print_perc(summ_brm$icc)` was attributable to differences between articles.

@@

# # Compute intra-class correlation
# var_total_reg <- var_experiment_reg + var_article_reg
# icc_reg <- var_article_reg / var_total_reg
# mean_qi(icc_reg)

# # Compute reduction in heterogeneity compared to the original meta-analysis
# 1 - mean(var_total_reg / var_total)
```

@@

(ref:fig2-caption) Effects of gender, age, and task. a Squares show the effect size (Hedges' g) for infants' mental rotation performance in each of the 59 individual experiments. Squares are color-coded according to the type of habituation task and the gender of the infants (blue = habituation task, all-female sample, yellow = habituation task, all-male sample, red = habituation task, mixed-gender sample, purple = violation of expectation [VoE] task, mixed-gender sample). Lines indicate the best-fit regression estimates according to a Bayesian three-level meta-regression model and gray ribbons indicate their corresponding 95% and 50% credible interval (CrI). b Fixed effect estimates obtained from the Bayesian three-level meta-regression model are depicted as black dots together with their 95% Crl (thin black lines) and 50% Crl (thick black lines). Gray curves indicate the posterior distribution for each effect.

To confirm the gender difference between males and females we set up another Bayesian meta-analysis, this time focusing on the looking-time contrasts between all-male and all-female groups reported within each article by the original authors.
Our additional analysis revealed a meta-analytic effect size of g = 0.20, 95% Crl [0.03, 0.38] (Supplementary Fig. 1) and a probability of this effect being greater than zero of 98.9%.
The heterogeneity of effect sizes was \$sigma\$2Experiment = 0.01, 95% CrI [0.00, 0.07], at the experiment level, and \$sigma\$2Article = 0.02, 95% CrI [0.00, 0.10] at the article level.
This means that 56.1% of the total variance can be attributed to differences between articles and 43.1% to differences within articles.

@@

```{r, fig4, include=TRUE, fig.width=6, fig.cap="(ref:fig4-caption)", out.width="50%"}
# Include the flowchart (created with MS PowerPoint)
knitr::include_graphics(here("results/figures/flowchart.pdf"))
```

@@

We entered the search terms ("mental rotation" OR "mental transformation" OR "spatial rotation" OR "spatial transformation" OR "spatial ability" OR "spatial skills") AND ("infant" OR "infants" OR "infanthood" OR "toddler" OR "toddlers" OR "toddlerhood" OR "child" OR "children" OR "childhood" OR "month" OR "months") into four online databases (APA PsycINFO, PubMed/MEDLINE, Scopus, ProQuest Dissertations & Theses Global).
All database queries were completed on December 6, 2021.
We configured the databases to check for article titles, abstracts, and keywords while applying no other filters or limits.
This yielded 2,616 articles in total, 1,954 of which remained after removing duplicate records (Figure 1).
We further identified 76 articles by screening the reference sections of previous reviews and meta-analyses on mental rotation and related skills (Frick et al., 2014; Johnson & Moore, 2020; Kubicek & Schwarzer, 2018; Lauer et al., 2019; Linn & Petersen, 1985; Moore & Johnson, 2020; Uttal et al., 2013; Voyer et al., 1995; Yang et al., 2020).
Of these, 34 had not been covered by the database search.
We also identified 94 articles by screening the reference sections of all articles that had been included after the first pass of the selection process.
Of these, 49 had not been covered by the database search.
We thus screened 2,037 unique articles in total.

@@

Two independent raters read the abstract and, if necessary, relevant sections of the full text to check if an article fulfilled the inclusion criteria.
Interrater agreement for the binary decision to include versus exclude an article was 98.5% ($Kappa$w [Cohen's weighted kappa] = 0.67, 95% CI [0.55, 0.78]).
Interrater agreement for the specific eligibility criteria were 88.2% ($Kappa$w = 0.72, 95% CI [0.40, 1.00]).
Cases where the two ratings diverged were resolved via discussion among all raters until a consensus was reached.
One article (Mash et al., 2007) was excluded because the authors used a unique mental rotation paradigm that was not comparable to that used in the other articles.
One article (Pedrett et al., 2020) was excluded because the average age of the infants (30.7 months) was almost twice as high as that of the next article (15.8 months), z = ???
compared to all articles.
We therefore decided to narrow our analysis from the first 3 years of life to the first 16 months of life.
This procedure led to a total of 23 articles being included in the meta-analysis (Fig. 4).

Many of these articles consisted of multiple experiments, e.g., using different variations of the mental rotation task or different subpopulations of infants.
We included all of these experiments in the meta-analysis and accounted for the dependencies between them by means of multilevel modeling with by-article random effects (see below).
However, we excluded experiments if there was insufficient information to compute a standardized effect size (see below).
Whenever an article reported separate effect sizes for boys and girls---or other subgroups like crawling and non-crawling infants)---but also an effect size combining these groups, we only included the combined effect size.
We dropped effect sizes that were clearly based on the same data but reported in different articles.
This procedure led to a total of 59 experiments being included in the meta-analysis (Supplementary Table 1).

@@

One outcome measure per experiment was converted into a standardized mean difference with small sample correction (Hedges' g) using the formulas provided in Table 1.
The sampling variance Var_g of the outcome measure g for each experiment was computed using the formula provided by Goulet-Pelletier & Cousineau (2018):

v = (df(df - 2)) \* ((2 \* (1 - r)) / n) \* (1 + g\^2 \* (n / (2 \* (1 - r)))) - (g\^2 / j\^2)
